---
title: "Amazon Reviews Sentiment Analysis"
author: "Nourhan Ghanima"
date: "2023-12-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Decoding Consumer Sentiments: Analyzing Amazon Reviews for Product Perception and User Satisfaction

Sentiment analysis is a technique that focuses on identifying the emotional tone or sentiment conveyed in a given piece of text.

It's particularly useful from a business perspective because it can provide important insight 
into customer opinions and preferences, which can then be used to make informed decisions about 
products or servies, after identifying potential areas of improvement. It also allows the monitoring of brand reputation, which is key to managing and enhancing brand perception. 


```{r echo = FALSE, message=FALSE}
library(tidyverse)
library(readr)
library(tm)
library(wordcloud)
library(syuzhet)
library(ggplot2)
```

```{r, include=FALSE, echo=FALSE}
reviews1 <- read_csv("/Users/nourhanghanima/downloads/amazon_reviews/1429_1.csv")
reviews2 <- read_csv("/Users/nourhanghanima/downloads/amazon_reviews/Datafiniti_Amazon_Consumer_Reviews_of_Amazon_Products_May19.csv")
reviews3 <- read_csv("/Users/nourhanghanima/downloads/Datafiniti_Amazon_Consumer_Reviews_of_Amazon_Products.csv")

reviews1 <- reviews1 |>
  select(id, name, asins, brand, categories, 
         manufacturer, reviews.numHelpful, 
         reviews.rating, reviews.text, reviews.title, reviews.username)

reviews2 <- reviews2 |>
  select(id, name, asins, brand, categories, 
                   manufacturer, reviews.numHelpful, 
                   reviews.rating, reviews.text, reviews.title, reviews.username)

reviews3 <- reviews3 |> 
  select(id, name, asins, brand, categories, 
         manufacturer, reviews.numHelpful, 
         reviews.rating, reviews.text, reviews.title, reviews.username)

amazon_reviews <- rbind(reviews1, reviews2, reviews3)


# changing Amazonbasics to Amazon Basics, to avoid duplication
amazon_reviews$brand <- sub("Amazonbasics", "AmazonBasics", amazon_reviews$brand)

```


The data for this project is taken from Kaggle, and includes user reviews left on a variety of products on Amazon.com.

After some basic data cleaning done on Excel and R with the tidyverse package, the descriptive statistics for the dataset are as follows:

```{r}
summary(amazon_reviews)
```

The data includes 67992 reviews, for a total of 7 brands, all owned by Amazon. 

## Which brand has the highest number of reviews?

```{r}
amazon_reviews |>
  group_by(brand) |>
  summarise(Count = n()) 
```

## Which brand has the highest average rating?

```{r, echo = FALSE}
#removing the brand Amazon Coco T because it has no star rating
average_rating <- amazon_reviews |>
  filter(brand != "Amazon Coco T") |>
  group_by(brand) |>
  summarise(average_rating = mean((reviews.rating), na.rm = TRUE))

```

From this plot showing average review ratings per brand, it's obvious that Amazon Digital Services has the lowest average rating. 

```{r, fig.width=12}
average_rating |>
  ggplot(aes(x = brand, y=average_rating)) +
  geom_col(fill= "blue") +
  xlab("Brand") +
  ylab ("Average Rating") +
  theme_bw()
```
```{r}
amazon_reviews |>
  filter(brand != "Amazon Coco T") |>
  group_by(brand) |>
  summarise(average_rating = mean((reviews.rating), na.rm = TRUE))
```
The average star rating for Amazon Fire TV is the highest. 

## Setting up sentiment analysis

The first step is to create a corpus.

```{r}
corpus <- iconv(amazon_reviews$reviews.text)
corpus <- Corpus(VectorSource(corpus))
```

I then transform the corpus to lowercase to avoid duplicates.

```{r, warning=FALSE}
corpus <- tm_map(corpus, tolower)
```

I then clean up the corpus by removing punctuation, numbers, and stopwords. 
```{r, warning = FALSE}
corpus <- tm_map(corpus, removeNumbers)
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, removeWords, stopwords("english"))
amazon_reviews_final <- corpus
```

Creating the term frequency matrix, which is a matrix of the frequency of each word in the reviews:
```{r}
tfm <- TermDocumentMatrix(amazon_reviews_final)
tfm <- as.matrix(tfm)
```

The first ten rows and five columns of the frequency matrix are displayed below.
```{r}
tfm[1:10, 1:5]
```

## Plot of word frequency

```{r}
w <- rowSums(tfm)
w <- subset(w, w >= 25)
barplot(w, las = 2, col = "blue")
```

I limit the words to words that have a frequency of 25 or higher. 

## The word cloud

```{r, echo = FALSE, warning = FALSE}
w <- sort(rowSums(tfm), decreasing = T)
wordcloud(words = names(w), 
          freq = w, 
          max.words = 40, 
          random.order = T,
          min.freq = 5,
          colors = brewer.pal(25, "Dark2"),
          scale = c(3, 0, .3)
          )
```

## Assigning sentiment scores for Amazon brand

```{r}
amazon_reviews_amazonbrand <- amazon_reviews |>
  filter(brand == "Amazon")
sentiment_data <- iconv(amazon_reviews_amazonbrand$reviews.text)
sentiment <- get_nrc_sentiment(sentiment_data)
sentiment[1:10, ]
```


## calculating the review wise score
```{r}
sentiment$score <- sentiment$positive - sentiment$negative
sentiment[1:30, ]

```

Most of the sentiment scores for the Amazon brand are positive. 
